Explore Vegetation Data

### 2015-12-21 explore_veg_data.Rmd
added a write out of canopy cover estimated for each plot via the spherical densiometer
Minor changes to enhance data summaries.

library(lubridate)
canopy_cover$year <- year(canopy_cover$date)
```

@@

Here I am adding a column that calculates the proportional coverage for each species based on the number of points they were observed in throughout the plot and the total number of points sampled in the plot.
```{r calculate overstory proportianal coverage}
mutate(veg_os_summ, pcov = ptct/totpts)
```

@@

unique(filter(veg_us, PlotID == "ann33", year == 2005)$UnderstorySp)
unique(filter(veg_us, PlotID == "ann33", year == 2011)$UnderstorySp)
veg_us$UnderstorySp[veg_us$PlotID == "ann33" & veg_us$year == 2011 & veg_us$UnderstorySp == "losp"] <- "lohi"
veg_us$UnderstorySp[veg_us$PlotID == "ann33" & veg_us$year == 2011 & veg_us$UnderstorySp == "syal"] <- "symo"
veg_us$UnderstorySp[veg_us$PlotID == "ann33" & veg_us$year == 2011 & veg_us$UnderstorySp == "roca"] <- "rogy"

@@

>>>>

library(lubridate)
canopy_cover$year <- year(canopy_cover$date)
canopy_cover <- select(canopy_cover, -date)
canopy_cover$plotid <- tolower(canopy_cover$plotid)
write.csv(canopy_cover, "analysis/data/plot_canopy_cover_dens.csv")
```

@@

Here I am adding a column that calculates the proportional coverage for each species based on the number of points they were observed in throughout the plot and the total number of points sampled in the plot.
```{r calculate overstory proportianal coverage at each point}
veg_os_summ <- mutate(veg_os_summ, pcov = ptct/totpts)
veg_os_summ
```

@@

unique(filter(veg_us, PlotID == "ann33", year == 2005)$UnderstorySp)
unique(filter(veg_us, PlotID == "ann33", year == 2011)$UnderstorySp)
filter(veg_us, UnderstorySp == "")
veg_us$UnderstorySp[veg_us$PlotID == "ann33" & veg_us$year == 2011 & veg_us$UnderstorySp == "losp"] <- "lohi"
veg_us$UnderstorySp[veg_us$PlotID == "ann33" & veg_us$year == 2011 & veg_us$UnderstorySp == "syal"] <- "symo"
veg_us$UnderstorySp[veg_us$PlotID == "ann33" & veg_us$year == 2011 & veg_us$UnderstorySp == "roca"] <- "rogy"

@@

Now I need to examine the data for alternate and non-host species DBH. These are the stems that are untagged and meet a 5cm DBH threshold.

```{r untagged species}
summary(alt_sp_dbh)
library(lubridate)
alt_sp_dbh$date <- as.Date(alt_sp_dbh$Date, format = "%m/%d/%Y")
filter(alt_sp_dbh, is.na(date)) # this should be 5/22/2012
alt_sp_dbh$Date[alt_sp_dbh$Date == ""] <- "5/22/2012 0:00:00"
alt_sp_dbh$date <- as.Date(alt_sp_dbh$Date, format = "%m/%d/%Y")
summary(alt_sp_dbh)
alt_sp_dbh$year <- year(alt_sp_dbh$date)
alt_sp_dbh <- alt_sp_dbh %>% select(-PlotVisit_PlotVisitID, -Date) %>% 
      rename(dbh = Measure)
unique(as.character(alt_sp_dbh$Species))

alt_sp_dbh$Species[alt_sp_dbh$Species == "PSME"] <- "psme"
alt_sp_dbh$Species <- tolower(alt_sp_dbh$Species)
summary(alt_sp_dbh)
unique(alt_sp_dbh$Species)
unique(alt_sp_dbh$Dead)
# alt_sp_dbh$Dead <- ifelse(alt_sp_dbh$Dead == 0, "no", "yes")
alt_sp_dbh <- as.tbl(alt_sp_dbh)

alt_sp_live <- filter(alt_sp_dbh, Dead == 0)
alt_sp_dead <- filter(alt_sp_dbh, Dead == 1)

alt_sp_summ <- left_join(
      # First data frame
      alt_sp_live %>% filter(year != 2012) %>% group_by(PlotID, Species, year) %>% 
      summarise(live_count = length(Species), live_avg_dbh = mean(dbh), live_tot_dbh = sum(dbh)),
      # Second data frame
      alt_sp_dead %>% filter(year != 2012) %>% group_by(PlotID, Species, year) %>% 
      summarise(dead_count = length(Species), dead_avg_dbh = mean(dbh), dead_tot_dbh = sum(dbh))
)
summary(alt_sp_summ)

write.csv(alt_sp_summ, "analysis/data/untagged_dbh_summary.csv")
```
